---
title: "FA22_virus_injections"
author: "Lorrie He"
date: "2022-09-13"
output: html_notebook
---

```{r setup, include=FALSE}

## load in packages
library(tidyverse)
library(dplyr)
library(plyr)
library(ggplot2)
#library(RColorBrewer)
#library(viridisLite)
#library(gridExtra)

# packages to read from google sheets
  # i haven't quite gotten this working yet...
# install.packages("googlesheets4")
#library(googlesheets4)

# last edited: 23 may 17 -- double check data filepaths before running!!!!

```

## read me & changelog

this is the cleaned up code (from manduca.R) for analysing the growth/etc data for the virus injections :)
see lab data dev/endpoint growth processing notes -- basically i need to re-do the analysis for this to have it match up to the field stuff (bc one tracks days after inject, other tracks days after hatch)

### changelog
reverse chronological

1/25/22:
* redoing round 4 plots --> make into bars for 1st committee mtg
  * copying from manduca.R
* to-do: 
  * fix file names LOL + add data into git repo
  * tbh i should put this into an R script. not an Rmd lol

9/25/22: 
* for now, work on the field analysis stuff before reworking the lab analysis.
* working on field dev



## dushay + beckage data
notes
  * values were obtained from plot in the paper via some webapp lol
  
  
```{r function defining}

testdata <- field_dev # create a test dataframe
  # 230125: 

# date-mo to julian
#to_julian <- function(column){
# column <-  as.Date(column, format = "%d-%b")
# df$j_date_c <- as.numeric(as.Date(df$date_c, format = "%j"))
#}
to_julian <- function(data, col){
  #data$col <- as.Date(data$col) #reformats dates
  data$j-{{col}} <- as.numeric(format(data$col, "%j"))
}

#to_julian(testdata,date.coll)


```


  
```{r d&b data}

dnb <- read.csv("~/OneDrive - University of North Carolina at Chapel Hill/Y1/virus/db1993_data.csv")
names(dnb) <- sub("^X", "", names(dnb)) # removes leading X's in column names

```


# 230125 figs for 1st comm mtg
basically regurgitate the R4 graphs in bars. use manduca.R a s a base


```{r lab data process and filter}

### read in data
lab <- read.csv("~/OneDrive - University of North Carolina at Chapel Hill/Documents/projects/ccbv_effects/data/lab_daily-growth.csv", header=T)

names(lab) <- sub("^X", "", names(lab)) # remove leading X's



### filter and remove bad samples, prep data to wrangle
lab_clean2 <- lab %>% select(-c(41:65, 37:39)) %>% # remove junk columns
  filter(!(unique == "R4-25-PBS-3")) %>% # remove super slow growers
  filter(!(unique == "R1-25-0.01-6")) %>%
  filter(!(para.type == "para" & fate == "wand")) %>% # removes unsuccesful paras
  filter(!(para.type == "para" & fate == "pup")) %>%
  filter(!(fate == "L" | fate == "na")) %>% #remove things that died as larvae (but keep dead pupa?)
  filter(!(para.type == "non")) #idk what to do w this

# keep R3 and R4 only
lab_new <- filter(lab_clean2, !(round == "R1" | round == "R2"))

lab_new <- lab_new[,-37] # remove "note" column at the end

# reshape df
lab_new.long <- lab_new %>% gather(key = "days.post.hatch", value = "mass", c(13:36))



### summary stats
stats.sub <- lab_new.long %>% group_by(para.type, heatshock, days.post.hatch) %>% 
  dplyr::summarise(avg = mean(mass, na.rm=TRUE), 
                   se = sd(mass, na.rm=TRUE)/length(mass),
                   n = n())

stats.sub$days.post.hatch <- as.numeric(stats.sub$days.post.hatch) # fixes days column lol



### plotting
lab_avgstats <- ggplot(data=stats.sub, aes(y=avg, x=days.post.hatch, 
                                           #color=as.factor(para.type), 
                                           color=as.factor(heatshock)
                                           ))

lab_avgstats + geom_point() + geom_errorbar(aes(ymin=avg-se, ymax=avg+se)) +
  facet_wrap(~para.type) +
  xlim(0, 25) + scale_color_brewer(palette = "Set1")


# taking out some irrelevant stuff for committee mtg
stats.sub2 <- filter(stats.sub, 
                     !(para.type == "0.50" | para.type == "2.00" ))

lab_avgstats2 <- ggplot(data=stats.sub2, aes(y=avg, x=days.post.hatch, 
                                           #color=as.factor(para.type), 
                                           color=as.factor(heatshock)
                                           ))

lab_avgstats2 + geom_line() + geom_errorbar(aes(ymin=avg-se, ymax=avg+se)) +
  facet_wrap(~para.type) +
  xlim(0, 25) + 
  scale_color_manual(values = c("25" = "blue", "40" = "red")) +
  #scale_color_brewer(palette = "Set1") +
  labs(x = "days after hatching", y = "average mass (g)", title = "average mass gain (lab)") +
  guides(color = guide_legend(title = "heatshock"))

```

```{r field data}

### read in data
field <- read.csv("data/fielding2.csv")


### fixing & wrangle prep
# fix column names
names(field) <- sub("^X", "", names(field))
#field_long <- field %>% gather(key = "date", value = "mass", c(9:50)) # gather all date info
  # for this code just keep using "long" df -- convert later

# convert "date...." columns to julian
long$date <- as.Date(long$date, format = "%d.%b")
long$jdate <- format(long$date, "%j")

long$date.inject <- as.Date(long$date.inject, format = "%d-%b")
long$jdate.inject <- format(long$date.inject, "%j")

long$jdate <- as.numeric(long$jdate) #jdate no longer julian
long$jdate.inject <- as.numeric(long$jdate.inject)

# create new "days since injection" column
long <- long %>% 
  #select(-c(10,13:14)) %>% #idky lol. probably old trouble columns
  mutate(days.since.inject = jdate-jdate.inject)



### stats
# some more cleaning to avoid breaking things
long$is.sup <- sub("^$", "no", long$is.sup) # fill empties w "no"
long <- long[,-1] # get rid of the ?s in the 1st column

field_MS <- long %>% # separates sups
  filter(species == "MS",
         days.since.inject >= 0,
         date.pmd == "",
         !(dose == "0.5" | dose == "uninj")) %>% # match up w/ lab graphs
  group_by(days.since.inject, temp, dose, is.sup, species) %>%
  dplyr::summarise(avg.mass = mean(mass, na.rm=TRUE),
                   se.mass = sd(mass, na.rm=TRUE)/length(mass),
                   n = n())




### plot
fieldplot <- field_MS %>% filter(is.sup == "no") %>% # remove sup for committee mtg so x-axis isnt cropped
  ggplot(aes(y=avg.mass, x=days.since.inject, 
             #color=as.factor(para.type), 
             color=as.factor(temp)
             ))

fieldplot + geom_line() + geom_errorbar(aes(ymin=avg.mass-se.mass, ymax=avg.mass+se.mass)) +
  facet_wrap(is.sup~dose) +
  xlim(0, 25) + 
  scale_color_manual(values = c("25" = "blue", "40" = "red")) +
  #scale_color_brewer(palette = "Set1") +
  labs(x = "days after injection", y = "average mass (g)", title = "average mass gain (field)") +
  guides(color = guide_legend(title = "heatshock"))



```



# 230125 PRE 1ST COMMITTEE MTG ERA
## lab data: development/endpoints

9/13/22 processing notes: 
  * i should actually process this more like the field data... go redo this lol
  * currently set up to be analysed as date since 3rd instar -- should be reworked to read by date since injection to better align with the field data

```{r initial pre-processing, warning=FALSE}

## read in data & some pre-processing
all_lab <- read.csv("~/OneDrive - University of North Carolina at Chapel Hill/Y1/virus/mar2022.csv")
  # 9/13/22: this is the longform data. go read in the short form and convert to long

all_lab <- all_lab %>% filter(days.since.3rd <= 15) #the numbers look bad after 13 days tbh bc of all the drops in weight
lab_wands <- read.csv("~/OneDrive - University of North Carolina at Chapel Hill/Y1/virus/wands.csv")

# make columns numeric (they dont read in this way for some reason)
all_lab$weight.g <- as.numeric(all_lab$weight.g)

lab_wands$days.died <- as.numeric(lab_wands$days.died)
lab_wands$days.em <- as.numeric(lab_wands$days.em)
lab_wands$days.pup <- as.numeric(lab_wands$days.pup)
lab_wands$days.wand <- as.numeric(lab_wands$days.wand)


```

```{r filtering/subset generation}

# to remove the bad injections, slow growers, wandering nat. paras, pre-mature deaths (in injected/NP groups)

lab_clean <- all_lab %>% filter(unique != "R1-para.d0-03", unique != "R1-para.d0-05", 
                            unique != "R1-para.d1-04",
                            unique != "R1-25-01", unique != "R1-25-02", unique != "R1-25-04", unique != "R1-25-09", unique != "R1-25-10",
                            unique != "R1-10-01", unique != "R1-10-02", unique != "R1-10-04", unique != "R1-10-07", unique != "R1-10-08", unique != "R1-10-09"  #keeping 10-10 in
                            ) %>%
                     filter((round == "R1" & treatment != "1") | round == "R2")


# 9/13/22 idk if i'll include those data subsets rn bc idk what i used them for LOL
#data.para <- data.rm2 %>% filter(unique != "R1-para.d0-03", unique != "R1-para.d0-05", 
#                             unique != "R1-para.d1-04", 
#                             treatment == "para")

#data.nopara <- data.rm2 %>% filter(treatment != "para")

#data.R1 <- data.rm2 %>% filter(round == "R1")

#data.R2 <- data.rm2 %>% filter(round == "R2")

#wands.nopara <- wands.rm %>% filter(treatment != "para")
#wands.para <- wands.rm %>% filter(treatment == "para")

```


```{r summary statistics}

summary.rm2 <- data.rm2 %>% group_by(treatment, days.since.3rd) %>% 
  summarise(avg = mean(weight.g, na.rm=TRUE),
            se = sd(weight.g, na.rm=TRUE)/length(weight.g),
            n = n())

summary.wands <- wands.nopara %>% group_by(treatment) %>%
  summarise(avg.wand.wt = mean(weight.wand, na.rm=TRUE),
            se.wand.wt = sd(weight.wand, na.rm=TRUE)/length(weight.wand),
            time.to.wand = days.wand-date.inj,
            avg.wand.time = mean(time.to.wand, na.rm=TRUE),
            se.wand.time = sd(time.to.wand, na.rm=TRUE)/length(time.to.wand),
            time.to.pup = days.pup-days.wand,
            avg.pup.time = mean(time.to.pup, na.rm=TRUE),
            se.pup.time = sd(time.to.pup, na.rm=TRUE)/length(time.to.pup),
            n = n())

```


```{r plots, echo=FALSE}

# sort this out again once you've cleaned up the lab processing

```


```{r lab data: development data part 2????}

# whatever the hell this is

#all_orig <- read.csv("~/OneDrive - University of North Carolina at Chapel Hill/Y1/virus/all_devdata.csv")
  # how is this different from ~/.../virus/mar2022.csv?

```

## lab data: daily growth

```{r read in lab}

growth_trimmed <- read.csv("~/OneDrive - University of North Carolina at Chapel Hill/Y1/virus/daily.csv", header=T)
names(growth_trimmed) <- sub("^X", "", names(growth_trimmed)) # remove leading X's

```

## field data: endpoints only

```{r read in field, message=FALSE}

field_dev <- read_csv("~/Desktop/working_csvs/field_dev.csv")

testdata <- testdata %>% 
  mutate(across(starts_with("mass.",), ~as.numeric(as.character(.))), # this one works
         across(starts_with("date.",), ~as.Date((.), format = "%d-%b")) # this doesnt tho lmfao
  )

## convert to julian dates

# something from indigo to try
# df %>%
#   mutate(
#     across(
#       .cols = date.inject:date.wander,
#       .fns = ~ as.numeric(as.Date(.x, format = '%d-%b'), format = '%j'),
#       .names = str_c('j', {.col}))
 
#field_dev$date.hatch <- as.Date(field_dev$date.hatch, format = "%d-%b")
#field_dev$j_date.hatch <- as.numeric(format(field_dev$hatch, "%j"))
#field_dev$date.4th <- as.Date(field_dev$date.4th, format = "%d-%b")
#field_dev$j_date.4th <- as.numeric(format(field_dev$date.4th, "%j"))
#field_dev$date.5th <- as.Date(field_dev$date.5th, format = "%d-%b")
#field_dev$j_date.5th <- as.numeric(format(field_dev$date.5th, "%j"))
#field_dev$date.6th <- as.Date(field_dev$date.6th, format = "%d-%b")
#field_dev$j_date.6th <- as.numeric(format(field_dev$date.6th, "%j"))
#field_dev$date.7th <- as.Date(field_dev$date.7th, format = "%d-%b")
#field_dev$j_date.7th <- as.numeric(format(field_dev$date.7th, "%j"))
field_dev$date.inject <- as.Date(field_dev$date.inject, format = "%d-%b")
field_dev$j_date.inject <- as.numeric(format(field_dev$inject, "%j"))
field_dev$date.pmd <- as.Date(field_dev$date.pmd, format = "%d-%b")
field_dev$j_date.pmd <- as.numeric(format(field_dev$date.pmd, "%j"))
field_dev$date.wander <- as.Date(field_dev$date.wander, format = "%d-%b")
field_dev$j_date.wander <- as.numeric(format(field_dev$date.wander, "%j"))
field_dev$date.pupa <- as.Date(field_dev$date.pupa, format = "%d-%b")
field_dev$j_date.pupa <- as.numeric(format(field_dev$date.pupa, "%j"))
field_dev$date.eclose <- as.Date(field_dev$date.eclose, format = "%d-%b")
field_dev$j_date.eclose <- as.numeric(format(field_dev$date.eclose, "%j"))

## doing stats
# stats_field.dev <- field_dev %>%      # this got wonky.. redoing this below
#     filter(species == "MS",
#          dose != "") %>%
#   group_by(temp, dose, is.sup) %>%
#   mutate(
#     time.towander = j_date.wander - j_date.inject,
#     time.topupate = j_date.pupa - j_date.inject,
#     time.toeclose = j_date.eclose - j_date.inject,
#     time.topmd = j_date.pmd - j_date.inject,
#     time.todeath = j_date.died - j_date.inject,
#   )

# avg masses and times to endpoints
endpoints_ms_field.dev <- field_dev %>%
  filter(species == "MS",               # filtering out "bad" individuals
         dose != "",
         source != "lab",
         is.na(ignore.reason)) %>%
  # mutate(                               # mutate adds onto new columns, summarise drops
  #   time.towander = j_date.wander - j_date.inject,
  #   time.topmd = j_date.pmd - j_date.inject) %>%
  group_by(temp, dose, is.sup) %>%
  dplyr::summarise(
    n = n(),
    avg_mass.wand = mean(mass.wander, na.rm=TRUE),
    avg_mass.pupa = mean(mass.pupa, na.rm=TRUE),
    avg_time.wand = mean(j_date.wander-j_date.inject),
    avg_time.pupa = mean(j_date.pupa-j_date.wander),
    avg_time.ec = mean(j_date.eclose-j_date.inject)
  )
  
# overview of individuals in expt (probably done!)
  # this would look nice as a stacked bar graph if i knew how to do that
  # consider when to separate MS and MQ...
overview_all_field.dev <- field_dev %>%
  filter(
    dose != "",
    species != "",
    source != "lab") %>%
  group_by(dose, temp, species) %>%
  dplyr::summarise(
    n = n(),
    n_pink = length(which(color == "pink")),
    n_pmd = length(which(outcome == "pmd")),
    n_wand = length(which(outcome == "wand")),
    n_wowe = length(which(outcome == "wowe")),
    n_em = length(which(outcome == "em")), 
    n_sup = length(which(is.sup == "yes")),
    pct_pmd = format(round(n_pmd/n*100, 1), nsmall=1),
    pct_wand = format(round(n_wand/n*100, 1), nsmall=1),
    pct_sup = format(round(n_sup/n*100, 1), nsmall=1),
    pct_wowe = format(round(n_wowe/n*100,1), nsmall=1),
    pct_em = format(round(n_wowe/n*100,1), nsmall=1)
  )




```

